A Study on Neural Models for Target-Based Computer-Assisted Musical Orchestration

### 2020-08-21 paper/aimusic2020/paper.tex
biblio and typos

% the name(s) of the author(s) follow(s) next
%
% NB: Chinese authors should write their first names(s) in front of
% their surnames. This ensures that the names appear correctly in
% the running heads and the author index.
%
\author{Carmine Emanuele Cella\inst{1}\and Luke Dzwonczyk\inst{1}\and Alejandro Saldarriaga-Fuertes\inst{1}\and Hongfu Liu\inst{1}\and Helene-Camille Crayencour\inst{2}}
%
% if the names of the authors are too long for the running head, please use the format: AuthorA et al.
\authorrunning{Carmine Emanuele Cella et al.}

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published
\institute{University of California, Berkeley \and Centrale Sup\'elec, L2S, Univ. Paris-Saclay, CNRS \\ \email{carmine.cella@berkeley.edu, dz.luke@berkeley.edu }}

@@

The development of computational tools to assist and inspire the musical composition process constitutes an import research area known as \emph{Computer-Assisted Composition (CAC)} \cite{FerVic2013, Ari2005}. Within CAC, target-based computer-assisted orchestration is an important example of how machine learning can be used for {enhancing} and {assisting} music creativity \cite{Maresz2003}.

@@

The second and deeper model that we trained as classifier was the well known deep residual network \emph{ResNet}. Specifically, we used 18-layer ResNet, which allows information to pass directly from the input to the final layer of each block. To make the model more suitable to our problem, we decided to reduce the number of parameters and to reset the output channel numbers of each block with 32, 64, 32, 32 respectively.

@@

During training, the loss function used to optimize the inner parameters of the model was binary cross entropy, as it is the common choice for multiclass multilabel classification frameworks. However, the value of the loss function alone is difficult to interpret.
%For this reason we created a complementary function $f$ to be used for evaluation only. This function compares a vector of expected outputs $\overline{X}$ with the estimated output from the model $\hat{X}$ by using the following function
%
%\begin{equation}
%f(\overline{X}, \hat{X}) = \frac{1}{N}<\overline{X}, M_N(\hat{X})>
%\label{eval}
%\end{equation}
%\begin{equation}
%M_N(\hat{X})_i = \left\{\begin{array}{ll}
%1 \text{ if } i \in I_N(\hat{X})\\
%0 \text{ otherwise}
%\end{array}\right.
%\label{NMax}
%\end{equation}
%and $I_N(\hat{X})$ is the set of indices containing the $N$ first maximums of the vector $\hat{X}$. More specifically, the function $M_N(\hat{X})$ takes as an input a vector of probabilities and outputs a vector where only the positions of the $N$ first maxima are set to 1. This new vector would be the orchestration of $N$ instruments given by the model. Thus, the function $f$ simply outputs the proportion of the estimated orchestration that matches the expected one.
For this reason, the evaluation was done by comparing the proportion of estimated orchestration samples, chosen among the samples that output the highest probability, that matched the expected orchestration.

@@

Another important extension would be to create a more powerful embedding spaces for the target and combinations. In \cite{Gillick19} the authors propose to use LSTM-based models to predict the embedding features for the combinations used during the optimisation process in assisted orchestration. We believe that by combining their prediction model with our classification models could generate more faithful representations and improve the overall quality of generated orchestrations.

>>>>

% the name(s) of the author(s) follow(s) next
%
% NB: Chinese authors should write their first names(s) in front of
% their surnames. This ensures that the names appear correctly in
% the running heads and the author index.
%
\author{Carmine Emanuele Cella\inst{1}\and Luke Dzwonczyk\inst{1}\and Alejandro Saldarriaga-Fuertes\inst{1}\and Hongfu Liu\inst{1}\and H\'el\`ene-Camille Crayencour\inst{2}}
%
% if the names of the authors are too long for the running head, please use the format: AuthorA et al.
\authorrunning{Carmine Emanuele Cella et al.}

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published
\institute{CNMAT - University of California, Berkeley \and Centrale Sup\'elec, L2S, Univ. Paris-Saclay, CNRS \\ \email{carmine.cella@berkeley.edu, dz.luke@berkeley.edu }}

@@

The development of computational tools to assist and inspire the musical composition process constitutes an important research area known as \emph{Computer-Assisted Composition (CAC)} \cite{FerVic2013, Ari2005}. Within CAC, target-based computer-assisted orchestration is a compelling case of how machine learning can be used for {enhancing} and {assisting} music creativity \cite{Maresz2003}.

@@

The second and deeper model that we trained as classifier was the well known deep residual network \emph{ResNet} \cite{He15}. Specifically, we used 18-layer ResNet, which allows information to pass directly from the input to the final layer of each block. To make the model more suitable to our problem, we decided to reduce the number of parameters and to reset the output channel numbers of each block with 32, 64, 32, 32 respectively.

@@

During training, the loss function used to optimize the inner parameters of the model was binary cross entropy, as it is the common choice for multiclass multilabel classification frameworks. However, the value of the loss function alone is difficult to interpret.

%For this reason we created a complementary function $f$ to be used for evaluation only. This function compares a vector of expected outputs $\overline{X}$ with the estimated output from the model $\hat{X}$ by using the following function
%
%\begin{equation}
%f(\overline{X}, \hat{X}) = \frac{1}{N}<\overline{X}, M_N(\hat{X})>
%\label{eval}
%\end{equation}
%\begin{equation}
%M_N(\hat{X})_i = \left\{\begin{array}{ll}
%1 \text{ if } i \in I_N(\hat{X})\\
%0 \text{ otherwise}
%\end{array}\right.
%\label{NMax}
%\end{equation}
%and $I_N(\hat{X})$ is the set of indices containing the $N$ first maximums of the vector $\hat{X}$. More specifically, the function $M_N(\hat{X})$ takes as an input a vector of probabilities and outputs a vector where only the positions of the $N$ first maxima are set to 1. This new vector would be the orchestration of $N$ instruments given by the model. Thus, the function $f$ simply outputs the proportion of the estimated orchestration that matches the expected one.
For this reason, the evaluation was done by comparing the proportion of estimated orchestration samples, chosen among the samples that output the highest probability, that matched the expected orchestration.

@@

Another important extension would be to create a more powerful embedding spaces for the target and combinations. In \cite{Gillick19} the authors propose to use LSTM-based models to predict the embedding features for the combinations used during the optimisation process in assisted orchestration. We believe that by combining their prediction model with our classification models we could generate more faithful representations and improve the overall quality of generated orchestrations.
